---
title: "config package for R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setenv(R_CONFIG_NAME = "default")
```

The **config** package makes it easy to manage environment specific configuration values (e.g. to use distinct values for development, testing, and production environments). 

## Basic Usage

Configurations are defined using a [YAML](http://www.yaml.org/about.html) text file and are stored (by default) in a file named `config.yml`. Configuration files include default values as well as values for arbitrary other named configurations, for example:

```yaml
default:
  trials: 5
  dataset: "data-sampled.csv"
  
production:
  trials: 30
  dataset: "data.csv"
```

To read values from this configuration file you call the `config::get` function, which returns a list containing all of the values for the currently active configuration:

```{r}
config <- config::get()
config$trials
```

You can also read a single value from the configuration as follows:

```{r}
trials <- config::get("trials")
```

The `get` function takes a `config` argument which determines which configuration to read values from. By default, `config` is read from the `R_CONFIG_NAME` environment variable. This variable is in turn typically set within a site sitewide `Renviron` or `Rprofile` (see [R Startup](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Startup.html) for details on these files). 

```{r}
Sys.setenv(R_CONFIG_NAME = "production")
config::get("trials")
```

```{r, include=FALSE}
Sys.setenv(R_CONFIG_NAME = "default")
```

<pre>
some test output code
</pre>



- read from an alternate configuration (via env variable)

- executing r code

- alternate config file

- inherit from default automatically
- inherit from other configuration

- "local" config file




